---
title: "Econometrics II - Assignment 3"
author: "Uncensored sloths"
date: "20 Jan 2022"
output:
  pdf_document:
    includes:
      in_header: "preamble.tex"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RCT)
library(dplyr)
library(magrittr)
library(lmtest)
library(stargazer)
library(lmtest)
library(sandwich)
```

# Question 1

(i) Compute for each of the colors the treatment effect.

(ii) Compute the average treatment effect in the full population.

(iii) Compute the average treatment effect on the treated.

(iv) Give an example where the average treatment effect on the treated would be more useful to consider than the average treatment effect, and explain why.

# Question 2

```{r}
# Load data
data <- read.csv("assignment3.csv")

data$important[data$important == "sport"] <- 0
data$important[data$important == "school"] <- 1
data$better[data$better == "Math"] <- 1
data$better[data$better == "Language"] <- 0
data$preferredhand[data$preferredhand == "Left"] <- 1
data$preferredhand[data$preferredhand == "Right"] <- 0

```

```{r}
data$treatment [data$treatment  == "Nothing"] <- 0
data$treatment [data$treatment  == "Candybar"] <- 2
data$treatment [data$treatment  == "Fruit"] <- 1

data$important[data$important == ""] <- NA

data %<>% 
      mutate_each(funs(if(is.character(.)) as.integer(.) else .))
```

(i) Compute the fraction of pupils in all three groups (control, fruit, and candy bar) that have the number correct and that are expected to lie. Show within a table whether pupil characteristics are balanced over the treatment groups.


```{r}
balance <- balance_table(data[, !names(data) %in% "id"], "treatment")
balance
```
For all three groups we can see that the average of correct numbers is above $\frac{1}{6}$ which is why calculate the fraction of pupils that lie using the given formula $\frac{x-\frac{1}{6}}{1-\frac{1}{6}}$:

```{r}
fraction_nothing <- (balance[2, 2]-(1/6))/(1- (1/6))
fraction_fruit <- (balance[2, 3]-(1/6))/(1- (1/6))
fraction_candybar <-(balance[2, 4]-(1/6))/(1- (1/6))
fractions <- cbind(fraction_nothing, fraction_fruit, fraction_candybar)
fractions
```

```{r}
data$treatment_fruit [data$treatment  == 0] <- 0
data$treatment_fruit [data$treatment  == 2] <- 0
data$treatment_fruit [data$treatment  == 1] <- 1
data$treatment_candy [data$treatment  == 0] <- 0
data$treatment_candy [data$treatment  == 2] <- 1
data$treatment_candy [data$treatment  == 1] <- 0
```


(ii) Use the linear probability model to regress the dummy variable for having the correct number on the dice on the assignment to the three treatment groups. Do you detect significant lying? Show how robust you results are to including additional covariates.

```{r}
model1 <- lm(correct ~ treatment_fruit + treatment_candy, data = data)
model1_robust <- coeftest(model1, vcov = vcovHC, type = "HC1")
model1_robust
```
they are not significantly different from $\alpha$ 

```{r}
t.test(subset(data, treatment == 0)$correct, y = NULL,
       alternative = c("greater"),
       mu = 1/6, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
```

```{r}
t.test(subset(data, treatment == 1)$correct, y = NULL,
       alternative = c("greater"),
       mu = 1/6, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
```

```{r}
t.test(subset(data, treatment == 2)$correct, y = NULL,
       alternative = c("greater"),
       mu = 1/6, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
```

```{r}
model2 <- lm(correct ~ treatment_fruit + treatment_candy + gender + grade +  important + better + preferredhand + siblings + youngestchild + oftenexpelled, data = data)
model2_robust <- coeftest(model2, vcov = vcovHC, type = "HC1")
model2_robust
```

```{r}
model3 <- lm(correct ~ treatment_fruit + treatment_candy + gender + better + preferredhand + siblings + youngestchild + oftenexpelled, data = data)
model3_robust <- coeftest(model3, vcov = vcovHC, type = "HC1")
model3_robust
```

```{r}
model4 <- lm(correct ~ treatment_fruit + treatment_candy + gender, data = data)
model4_robust <- coeftest(model4, vcov = vcovHC, type = "HC1")
model4_robust
```

We tried adding different covariates. However, except for gender no covariate was significant and additionally led to higher standard errors. The same happens if one includes combinations of covariates into the model as shown in the second and third model.

(iii) The researcher is interested in testing if boys and girls behave differently. Write down and estimate a model that can test if boys lie more than girls and whether boys and girls respond differently to incentives (e.g. treatment).

```{r}
model4_robust
```
In the model including gender as a regressors, we can see that gender is significant at $10\%$ and negative. As the gender dummy equals 1 for girlst, we can conclude that there is weakly significant effect and the girls tend to lie less than boys.

```{r}
model5 <- lm(correct ~ treatment_fruit + treatment_candy + gender*treatment_fruit + gender*treatment_candy, data = data)
model5_robust <- coeftest(model5, vcov = vcovHC, type = "HC1")
model5_robust
```



(iv) Estimate your preferred model specification (including pupil characteristics if this improves your model) and write down your main conclusions.
Prior to conducting the experiment, the researcher performed some power calculations.

treatments and gender, no interaction effect
children just always lie


(v) Given the sample size and the estimates you have obtained above, what would be the minimum detectable effect size of this experiment?

```{r}
# Fruit
k=4 
q=0.8
alpha=0.05

n <- length(data$treatment)
n1 <- length(data$treatment[data$treatment  == 1])

t_alpha <- qt(p=alpha/2, df=n-k, lower.tail=FALSE)
t_q <- qt(p=q, df=n-k, lower.tail=TRUE) # denk nochmal drüber nach
p1 = n1/n

#Residual Standard error (Like Standard Deviation)
k=length(model4$coefficients)-1 #Subtract one to ignore intercept
SSE=sum(model4$residuals**2)
m=length(model4$residuals)
sigma = sqrt(SSE/(m-(1+k))) #Residual Standard Error

MDE_fruit = (t_alpha - t_q)*sqrt(1/(p1*(1-p1)))*sqrt(sigma^2/n)

# Candybar

n2 <- length(data$treatment[data$treatment  == 2])
p2 = n2/n

MDE_candy = (t_alpha - t_q)*sqrt(1/(p2*(1-p2)))*sqrt(sigma^2/(n))

print(MDE_fruit)
print(MDE_candy)
```


(vi) Initially, the researcher expected that no pupil would lie if there is nothing at stake (control group) and that a quarter of the student lie for a candy bar. How large should the sample size of the experiment have been in that case?

```{r}
MDE_target = 0.25*(1 - (1/6))
samplesize = ((t_alpha - t_q)/MDE_target)^2*((sigma^2)/(p2*(1-p2)))
print(samplesize)
```

Note: same df used as above to not have to iteratively solve it (only margignal difference)

(vii) Students assigned to the control group provide counterfactuals to both the fruit and the candy bar treatment. Show that by assigning more students to the control group than to each treatment group the same MDE can be achieved with fewer students in the experiment.

```{r}
p1_new = p1+0.1
p2_new = p2+0.1
n_1 = ((t_alpha - t_q)/MDE_fruit)^2*(sigma^2)/(p1_new*(1-p1_new))
n_2 = ((t_alpha - t_q)/MDE_candy)^2*(sigma^2)/(p2_new*(1-p2_new))
samplesize = max(n_1,n_2)
print(samplesize)
```
??? Assignment wants to increase treatment, but function is minimal for p=0.5, which also makes no sense, since then no one is in the control so there must be a fundamental mistake in how we calculate the stuff, uff
